---
title: "Music Physiological Response Analysis"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: show
    theme: cosmo
date: "2025-02-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 10, 
  fig.height = 6,
  out.width = "100%"
)
```

## 1. Load Packages and Set Theme

```{r packages}
# Load necessary packages

library(tidyverse)
library(data.table)
library(lubridate)
library(lme4)      # For mixed linear models
library(lmerTest)  # For p-values in mixed models
library(sjPlot)    # For model visualization
library(ggplot2)   # For visualization
library(patchwork) # For combining plots
library(scales)    # For better axis formatting
library(psych)     # For descriptive statistics

# Set a consistent theme for all plots
theme_set(theme_minimal(base_size = 12) + 
          theme(legend.position = "bottom",
                panel.grid.minor = element_blank()))
```

## 3. Load and Process Downsampled Data

```{r}


# Load the downsampled data from CSV (replace with your actual file path)
continuous_data_1Hz <- read.csv("C:/Users/Elias/Documents/continuous_data_1Hz_avg.csv")

# Make sure all numeric columns are properly typed
numeric_cols <- c("Time", "RelativeTime", "BVP", "Corr", "Resp", "Skin", "Temp", "Trap", "Zygo")
continuous_data_1Hz[numeric_cols] <- lapply(continuous_data_1Hz[numeric_cols], as.numeric)

# Convert categorical variables to factors if needed
factor_cols <- c("Participant", "Session", "StimID", "StimOrder")
continuous_data_1Hz[factor_cols] <- lapply(continuous_data_1Hz[factor_cols], as.factor)

# Load post-stimulus data
post_stim_files <- list.files(
  path = "C:/Users/Elias/Downloads/5632210",
  pattern = glob2rx("PostStimulus_Numeric*.csv"), 
  recursive = TRUE,
  full.names = TRUE
)
# Function to read post-stimulus numeric data with improved error handling
read_post_stimulus <- function(file_path) {
  tryCatch({
    data <- read.csv(file_path, header = FALSE)
    
    # Ensure we have the expected number of columns
    if(ncol(data) >= 18) {
      colnames(data) <- c("ReportTime", "StartTime", "EndTime", "Interrupted", "StimOrder", 
                         "StimID", "ParticipantNumber", "Cough", "Sneeze", "Laugh", "Doze", 
                         "Yawn", "ShiverChills", "MoveToMusic", "Other", "Familiarity", 
                         "Liking", "Focus", "SelfConsciousness")[1:ncol(data)]
    } else {
      warning(paste("Post-stimulus file has unexpected column count:", file_path))
      return(NULL)
    }
    
    # Extract session from filename with regex
    file_name <- basename(file_path)
    session_match <- regexpr("Sess(\\d+)", file_name)
    participant_match <- regexpr("P(\\d+)", file_name)
    
    data$Session <- if(session_match > 0) 
      as.numeric(gsub("Sess", "", regmatches(file_name, session_match))) else NA
    data$Participant <- if(participant_match > 0) 
      as.numeric(gsub("P", "", regmatches(file_name, participant_match))) else NA
    
    # Convert numeric data to appropriate types
    numeric_cols <- c("Interrupted", "StimOrder", "StimID", "Cough", "Sneeze", 
                     "Laugh", "Doze", "Yawn", "ShiverChills", "MoveToMusic", "Other",
                     "Familiarity", "Liking", "Focus", "SelfConsciousness")
    numeric_cols <- numeric_cols[numeric_cols %in% names(data)]
    
    for(col in numeric_cols) {
      data[[col]] <- as.numeric(as.character(data[[col]]))
    }
    
    return(data)
  }, error = function(e) {
    warning(paste("Failed to process post-stimulus file:", file_path, "\nError:", e$message))
    return(NULL)
  })
}



cat("Processing", length(post_stim_files), "post-stimulus files...\n")
post_stimulus_list <- lapply(post_stim_files, read_post_stimulus)
post_stimulus_list <- post_stimulus_list[!sapply(post_stimulus_list, is.null)]
post_stimulus_data <- bind_rows(post_stimulus_list)
cat("Loaded", nrow(post_stimulus_data), "rows of post-stimulus data\n")

# Convert StimID to numeric before joining

continuous_data_1Hz <- continuous_data_1Hz %>%
  mutate(StimID = as.numeric(as.character(StimID)))

# Then create and join the metadata
stimulus_metadata <- tibble(
  StimID = c(101, 102, 103, 104, 105, 106, 107, 
             211, 212, 213, 214, 
             221, 222, 223, 224, 
             231, 232, 233, 234, 
             241, 242, 243, 244),
  Title = c("String Quartet Op. 131, II - III", "Stampede", "Basket", "1685/Bach", 
            "O Fortuna, Carmina Burana", "Le rosier de trois couleurs de roses", "Visiting Hours",
            "Timshel", "Reapers", "L'inverno (Winter): I. Allegro Non Molto", "Nanou 2",
            "Endless Love", "Dare You to Move", "Enchanted Suite", "Love Theme for Nata",
            "Always Be My Baby", "Hard In Da Paint", "Animals", "Europa (Earth's Cry Heaven's Smile)",
            "Ojos Color Sol", "It's Still Rock and Roll to Me", "Rodeo: Hoe-Down", "25-1-14-14"),
  Artist = c("Artemis Quartet/L. v. Beethoven", "The Quantic Soul Orchestra/Will Holland", 
             "Dan Mangan", "Nosaj Thing", "Orch. Symph. Montreal/Carl Orff", 
             "STRADA/Anon", "Shane Koyczan", "Mumford & Sons", "Muse", 
             "Joshua Bell & Academy of St. Martin in the Fields/A. Vivaldi", "Aphex Twin",
             "Luther Vandross/L. Richie", "Switchfoot/J. Foreman", "Alan Menken", "Ennio Morricone",
             "Mariah Carey/M. Carey & J. Dupri & M. Seal", "Waka Flocka Flame", "Martin Garrix", 
             "Santana/C. Santana & T. Coster", "Calle 13/E. Cabra", "Billy Joel", 
             "Leonard Slatkin & Saint Louis Symphony Orchestra/A. Copland", "Pierre Lapointe"),
  Genre = c("Classical", "Funk", "Singer/Songwriter", "Dubstep/Glitch-hop", 
            "20th C. Classical", "French Canadian Folksong", "Spoken word", 
            "Folk", "Rock", "Baroque", "Piano", 
            "R & B", "Pop", "Soundtrack", "Soundtrack", 
            "Pop", "Hip-hop/Rap", "Electronic Dance Music", "Rock", 
            "Latin Urban", "Rock'n'Roll", "Classical", "French Pop"),
  Duration_s = c(215, 188, 232, 168, 159, 146, 250, 
                 173, 360, 204, 204, 
                 258, 247, 272, 247, 
                 258, 246, 304, 306, 
                 217, 177, 261, 142)
) %>%
  mutate(
    ExperimentSelected = substr(as.character(StimID), 1, 1) == "1",
    ParticipantSelected = substr(as.character(StimID), 1, 1) == "2",
    # Add broad genre categories
    GenreCategory = case_when(
      Genre %in% c("Classical", "Baroque", "20th C. Classical", "Piano") ~ "Classical",
      Genre %in% c("Rock", "Rock'n'Roll") ~ "Rock",
      Genre %in% c("Pop", "R & B", "Singer/Songwriter", "Folk") ~ "Pop/Folk",
      Genre %in% c("Hip-hop/Rap", "Electronic Dance Music", "Dubstep/Glitch-hop", "Funk") ~ "Electronic/Urban",
      Genre %in% c("Soundtrack") ~ "Soundtrack",
      TRUE ~ "Other"
    )
  )

# Then join
continuous_data_1Hz <- continuous_data_1Hz %>%
  left_join(stimulus_metadata, by = "StimID")

```
Plot
```{r}
# Visualization to confirm no extreme outliers remain
cleaned_boxplots <- continuous_data_1Hz %>%
  pivot_longer(cols = c(BVP, Corr, Resp, Skin, Temp, Trap, Zygo), 
               names_to = "Measure", values_to = "Value") %>%
  ggplot(aes(x = Measure, y = Value)) +
  geom_boxplot() +
  facet_wrap(~Measure, scales = "free_y") +
  labs(title = "Distributions of Cleaned Physiological Measures",
       y = "Value") +
  theme(axis.text.x = element_blank())

print(cleaned_boxplots)
continuous_data_1Hz
```

### 4.3 Physiological Signal Feature Extraction

```{r feature_extraction}
# Calculate more advanced physiological features

physio_features <- continuous_data_1Hz %>%
  # Group by each stimulus presentation
  group_by(Participant, Session, StimOrder, StimID) %>%
  summarize(
    # Time domain features
    mean_BVP = mean(BVP, na.rm = TRUE),
    sd_BVP = sd(BVP, na.rm = TRUE),
    
    # BVP features - calculate heart rate variability approximation
    # First 20 seconds often contains adaptation, so we'll use data after that
    # Assuming 100Hz sampling rate
    hr_var = if(n() > 3000) sd(BVP[RelativeTime >= 20], na.rm = TRUE) else NA,
    
    # Skin conductance features
    mean_Skin = mean(Skin, na.rm = TRUE),
    sd_Skin = sd(Skin, na.rm = TRUE),
    # Calculate skin conductance response frequency (SCR)
    # Simple approximation: count instances where skin conductance increases by 0.05 units within 3 seconds
    scr_count = sum(diff(Skin) > 0.05, na.rm = TRUE),
    scr_rate = scr_count / (max(RelativeTime, na.rm = TRUE) / 60), # per minute
    
    # Respiratory features
    mean_Resp = mean(Resp, na.rm = TRUE),
    sd_Resp = sd(Resp, na.rm = TRUE),
    # Improved respiration rate calculation using zero crossings with filtering
    resp_rate = sum(diff(sign(Resp - mean(Resp, na.rm = TRUE))) != 0, na.rm = TRUE) / 
               (2 * max(RelativeTime, na.rm = TRUE) / 60), # divide by 2 to count full cycles
    
    # EMG features (facial expressions)
    mean_Corr = mean(Corr, na.rm = TRUE), # Corrugator - frowning
    sd_Corr = sd(Corr, na.rm = TRUE),
    mean_Zygo = mean(Zygo, na.rm = TRUE), # Zygomaticus - smiling
    sd_Zygo = sd(Zygo, na.rm = TRUE),
    mean_Trap = mean(Trap, na.rm = TRUE), # Trapezius - tension
    sd_Trap = sd(Trap, na.rm = TRUE),
    
    # Ratio of positive to negative facial expression (smiling vs frowning)
    smile_frown_ratio = mean_Zygo / mean_Corr,
    
    # Stimulus duration
    stim_duration = max(RelativeTime, na.rm = TRUE),
    
    # Calculate emotional metrics if available
    has_emotion_data = "EmotionalArousal" %in% names(.) & "EmotionalValence" %in% names(.),
    mean_arousal = if("EmotionalArousal" %in% names(.)) 
                      mean(EmotionalArousal, na.rm = TRUE) else NA,
    mean_valence = if("EmotionalValence" %in% names(.)) 
                      mean(EmotionalValence, na.rm = TRUE) else NA,
    
    # Calculate time trends - early vs. late response
    # Early: first third, Late: last third
    early_Skin = if(n() > 300) mean(Skin[RelativeTime <= quantile(RelativeTime, 0.33)], na.rm = TRUE) else NA,
    late_Skin = if(n() > 300) mean(Skin[RelativeTime >= quantile(RelativeTime, 0.67)], na.rm = TRUE) else NA,
    skin_change = late_Skin - early_Skin,
    
    early_Zygo = if(n() > 300) mean(Zygo[RelativeTime <= quantile(RelativeTime, 0.33)], na.rm = TRUE) else NA,
    late_Zygo = if(n() > 300) mean(Zygo[RelativeTime >= quantile(RelativeTime, 0.67)], na.rm = TRUE) else NA,
    zygo_change = late_Zygo - early_Zygo,
    
    # Sample count
    n_samples = n()
  )
```

### 4.4 Create Analysis Dataset

```{r analysis_dataset}
# Create analysis dataset by joining feature data with post-stimulus data
analysis_data <- physio_features %>%
  left_join(post_stimulus_data, by = c("Participant", "Session", "StimOrder", "StimID")) %>%
  left_join(stimulus_metadata, by = "StimID") %>%
  # Convert binary responses to logical for better interpretation
  mutate(across(c(Cough, Sneeze, Laugh, Doze, Yawn, ShiverChills, MoveToMusic, Other, 
                 ExperimentSelected, ParticipantSelected), 
                ~as.logical(as.numeric(.)))) %>%
  # Scale emotion ratings to 0-100 for interpretability
  mutate(
    Familiarity_pct = Familiarity / 1.27,
    Liking_pct = Liking / 1.27,
    Focus_pct = Focus / 1.27,
    SelfConsciousness_pct = SelfConsciousness / 1.27
  ) %>%
  # Final cleaning and quality flags
  mutate(
    data_quality_flag = case_when(
      Interrupted == 1 ~ "interrupted",
      is.na(resp_rate) | resp_rate < 8 | resp_rate > 25 ~ "abnormal_resp_rate",
      is.na(mean_BVP) | is.na(mean_Skin) ~ "missing_key_data",
      n_samples < 1000 ~ "insufficient_data",
      TRUE ~ "ok"
    ),
    # Create musical arousal proxy based on genre
    musical_arousal_proxy = case_when(
      Genre %in% c("Hip-hop/Rap", "Electronic Dance Music", "Rock", "Dubstep/Glitch-hop") ~ "high",
      Genre %in% c("Classical", "Baroque", "Soundtrack", "French Canadian Folksong", "Piano") ~ "low",
      TRUE ~ "medium"
    )
  )

# Check if we have any recordings with quality issues
quality_summary <- analysis_data %>%
  count(data_quality_flag) %>%
  mutate(percentage = n / sum(n) * 100)

print(quality_summary)
```

## 5. Exploratory Data Analysis

### 5.1 Overview of Dataset

```{r data_overview}
# Data overview - how many stimuli, sessions, etc.
data_counts <- analysis_data %>%
  summarize(
    n_sessions = n_distinct(Session),
    n_stimuli = n_distinct(StimID),
    n_observations = n(),
    n_experiment_selected = sum(ExperimentSelected, na.rm = TRUE),
    n_participant_selected = sum(ParticipantSelected, na.rm = TRUE)
  )

print(data_counts)

# Distribution of stimuli by genre
genre_distribution <- analysis_data %>%
  group_by(Genre) %>%
  summarize(
    count = n(),
    mean_liking = mean(Liking_pct, na.rm = TRUE),
    mean_familiarity = mean(Familiarity_pct, na.rm = TRUE)
  ) %>%
  arrange(desc(count))

print(genre_distribution)

# Visualize distribution of stimuli by genre with liking scores
genre_plot <- ggplot(genre_distribution, aes(x = reorder(Genre, -count), y = count, fill = mean_liking)) +
  geom_col() +
  scale_fill_viridis_c(option = "plasma", name = "Mean Liking (%)") +
  labs(title = "Distribution of Stimuli by Genre with Liking Scores",
       x = "Genre", y = "Count") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

print(genre_plot)
```

### 5.2 Descriptive Statistics for Key Variables

```{r descriptive_stats}
# Descriptive statistics for all numeric variables
descriptive_stats <- analysis_data %>%
  select(where(is.numeric), -ends_with("Time"), -contains("ID"), -Session, -Participant) %>%
  select(-Interrupted, -starts_with("Stim")) %>%
  psych::describe() %>%
  select(n, mean, sd, median, min, max, skew, kurtosis)

print(descriptive_stats)

# Compare physiological responses across genre categories
physio_by_genre <- analysis_data %>%
  group_by(GenreCategory) %>%
  summarize(
    mean_skin = mean(mean_Skin, na.rm = TRUE),
    sd_skin = sd(mean_Skin, na.rm = TRUE),
    mean_scr_rate = mean(scr_rate, na.rm = TRUE),
    mean_resp_rate = mean(resp_rate, na.rm = TRUE),
    mean_smile = mean(mean_Zygo, na.rm = TRUE),
    mean_frown = mean(mean_Corr, na.rm = TRUE),
    mean_liking = mean(Liking_pct, na.rm = TRUE),
    n = n()
  )

print(physio_by_genre)

# Visualize physiological responses by genre
physio_genre_plot <- physio_by_genre %>%
  pivot_longer(cols = starts_with("mean_"), names_to = "measure", values_to = "value") %>%
  ggplot(aes(x = GenreCategory, y = value, fill = GenreCategory)) +
  geom_col() +
  facet_wrap(~measure, scales = "free_y") +
  labs(title = "Physiological Responses by Genre Category",
       x = "Genre Category", y = "Value") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none")

print(physio_genre_plot)
```

### 5.3 Relationship between Subjective and Physiological Responses

```{r subjective_physio_relationship}
# Create correlation matrix for key variables
correlation_vars <- analysis_data %>%
  filter(data_quality_flag == "ok") %>%
  select(mean_Skin, scr_rate, mean_Zygo, mean_Corr, smile_frown_ratio, 
         Liking_pct, Familiarity_pct, Focus_pct, SelfConsciousness_pct)

correlation_matrix <- cor(correlation_vars, use = "pairwise.complete.obs")
corrplot::corrplot(correlation_matrix, method = "circle", type = "upper", 
                   tl.col = "black", tl.srt = 45, addCoef.col = "black", 
                   number.cex = 0.7)

# Explore relationship between liking and physiological responses
liking_plots <- analysis_data %>%
  filter(data_quality_flag == "ok") %>%
  select(Liking_pct, mean_Skin, mean_Zygo, mean_Corr, smile_frown_ratio) %>%
  pivot_longer(cols = c(mean_Skin, mean_Zygo, mean_Corr, smile_frown_ratio), 
               names_to = "measure", values_to = "value") %>%
  ggplot(aes(x = Liking_pct, y = value)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "loess", se = TRUE) +
  facet_wrap(~measure, scales = "free_y") +
  labs(title = "Relationship between Liking and Physiological Responses",
       x = "Liking (%)", y = "Value")

print(liking_plots)

# Visualize emotional reactions across sessions
emotion_by_session <- analysis_data %>%
  group_by(Session) %>%
  summarize(
    mean_liking = mean(Liking_pct, na.rm = TRUE),
    mean_familiarity = mean(Familiarity_pct, na.rm = TRUE),
    mean_focus = mean(Focus_pct, na.rm = TRUE),
    mean_self_consciousness = mean(SelfConsciousness_pct, na.rm = TRUE),
    mean_skin = mean(mean_Skin, na.rm = TRUE),
    mean_zygo = mean(mean_Zygo, na.rm = TRUE)
  ) %>%
  pivot_longer(cols = starts_with("mean_"), names_to = "measure", values_to = "value")

session_emotion_plot <- ggplot(emotion_by_session, aes(x = Session, y = value, group = measure, color = measure)) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  facet_wrap(~measure, scales = "free_y") +
  labs(title = "Changes in Emotional and Physiological Responses Across Sessions",
       x = "Session Number", y = "Value") +
  theme(legend.position = "none")

print(session_emotion_plot)
```

### 5.4 Response Patterns for Familiar vs. Unfamiliar Music

```{r familiar_unfamiliar}
# Create binary familiarity variable for analysis
analysis_data <- analysis_data %>%
  mutate(
    is_familiar = Familiarity_pct >= 50,
    is_liked = Liking_pct >= 50
  )

# Compare responses to familiar vs. unfamiliar music
familiarity_comparison <- analysis_data %>%
  group_by(is_familiar) %>%
  summarize(
    n = n(),
    mean_liking = mean(Liking_pct, na.rm = TRUE),
    mean_focus = mean(Focus_pct, na.rm = TRUE),
    mean_self_consciousness = mean(SelfConsciousness_pct, na.rm = TRUE),
    mean_skin = mean(mean_Skin, na.rm = TRUE),
    mean_zygo = mean(mean_Zygo, na.rm = TRUE),
    mean_corr = mean(mean_Corr, na.rm = TRUE),
    smile_ratio = mean(smile_frown_ratio, na.rm = TRUE),
    laughed_pct = mean(Laugh, na.rm = TRUE) * 100,
    moved_to_music_pct = mean(MoveToMusic, na.rm = TRUE) * 100,
  )

print(familiarity_comparison)

# Visualize key differences between familiar and unfamiliar music
familiarity_plots <- analysis_data %>%
  pivot_longer(cols = c(mean_Skin, mean_Zygo, mean_Corr, smile_frown_ratio), 
               names_to = "measure", values_to = "value") %>%
  ggplot(aes(x = is_familiar, y = value, fill = is_
